Overlapping Batches for the Assessment of Solution Quality in Stochastic Programs

### 2014-07-05 Referee1_Response.tex

\setcounter{MaxMatrixCols}{10}

\setlength{\textheight}{9in}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{-0.5in}
\setlength{\oddsidemargin}{0.0in}

\newcommand{\noi}{\noindent}

% Definitions for overlapping batches
\newcommand{\gb}{\bar{G}}
\newcommand{\gbb}{\bar{\gb}}
\newcommand{\db}{\bar{D}}
\newcommand{\dbb}{\bar{\db}}

@@

\noi 
Based on your and the associate editor's comments, we studied the computational efficiency of the OMRP estimators. 
The results of this study are presented in a new section, Section 5.4 (Computational Efficiency).
%We studied the computational efficiency of the OMRP estimators, based on your and the associate editor's comments, and provided a summary of our results in a new section, Section 5.4 (Computational Efficiency). 
We first approached efficiency by the reciprocal of variance multiplied by the computation time, per associate editor's suggestion. 
Maximizing the computational efficiency defined in this manner provided guidelines on selecting the degree of overlap. 
We then looked at how OMRP compared to increasing the number batches in MRP.
We examined computational efficiency both theoretically and empirically. 
For empirical evaluation, we used the newsvendor problem with insertion sort, based on your suggestions. 
We found that when effective warm-starting is available, overlapping can lead to more efficient estimators than (nonverlapping) MRP, both when the direct nonoverlapping counterpart is used or when the number of batches in MRP is increased to match the variance of the overlapping counterpart.  
All these results are presented in Section 5.4 (Computational Efficiency). 
Thank you for this suggestion. 
We believe the study of computational efficiency significantly improved the paper. 
\medskip

\begin{itemize}
\item[\textbf{M2.}] \textit{The choice of the lower bound that is embedded in equation (8) is interesting. 
(The upper bound in equation (8) is completely natural as the sample mean of the $n_l$ realizations of the function evaluated at $\hat{x}$.) 
It seems that there are also other natural choices that could be made here for the lower bound. 
There is some discussion of this choice in the text that follows equations (7)-(9), but can the authors say more? For example, do the consistency and asymptotic validity results that follow hinge on this choice? Or, do those results also hold for other choices, like forming $\gbb$ as a sample
mean of the underlying $\gb$ values. 
We know that the individual gap estimators in (7) are nonnegative, and if we formed $\gbb$ as a sample mean of $\gb$ values, it would be nonnegative. 
Is the $\gbb$, as specified in (8), nonnegative? 
Have the authors computationally explored other choices here? 
Any insight in this vein would be valuable to the reader in understanding the choice made in equation (8).}
\end{itemize}

\noi
You are right; we did not clearly motivate this choice of $\gbb$.  
We added explanations after the definition of the OMRP estimators, motivating this choice and relating it to the theoretical results presented in the paper.\smallskip

\noi 
As you pointed out for the upper bound in (8), this $\gbb$ was chosen for each observation to have equal weight when the sampling problems yield optimal solutions $x_j^* = x^*, \ \forall j$. 
This fact is used in Lemmas 1 and 2, which are then used---directly or indirectly---in proving Theorems 1--3. 
When the alternative $\gbb$, formed as a sample mean of $\gb$ values, is used, theoretical results presented in the paper can be recovered by appropriately adjusting the coefficients of the $\dbb$ terms.
Notice that these results establish convergence of  $\gbb$, $VG$, etc.\ to their nonoptimized counterparts $\dbb$, $VD$, etc.
When the coefficients of $\dbb$ are adjusted, however, $\dbb$ and $VD$ are {\it not exactly} the same as the overlapping estimators of simulation output analysis. 
Because our results depend on the asymptotic properties of the overlapping estimators already established in the simulation literature, we did not want to make any changes. 
It may be possible to recover the same results but we did not pursue it here. 
We actually performed computations using $\gbb$ formed as a sample mean of $\gb$ values and the results were essentially the same. 
However, due to subtle theoretical reasons explained above, we decided not to include or discuss this choice. 
\medskip

@@

\begin{itemize}
\item[18.] \textit{p18, l57: The authors speak of coverage probabilities for DB1, but how can this be done without knowing an optimal solution to DB1?}
\end{itemize}

\noindent 
We used a Latin hypercube sample of size 50,000 estimate $\mu_{\hat{x}}$ for DB1, which was then used to estimate the coverage probability of the OMRP confidence intervals for this test problem.
We explain this at the end of Section 5.1 (Test Problems). 
We also changed the wording in Section 5.3 (Results of Experiments) to make it clear that the coverage probability is estimated for DB1. 
\medskip

\begin{itemize}
\item[19.] \textit{p19: It may be possible to expand on the last paragraph of Section 5, in giving guidance on selecting the degree of overlap, depending on the authors' response to comment M1 above.}
\end{itemize}

\noindent 
Thank you for this suggestion. 
Based on our study of computational efficiency of the OMRP estimators, we  provided an updated and expanded guidance on selecting the degree of overlap. 
This guidance appears in the last paragraph of \S 5.4 (Computational Efficiency).

### 2014-07-05 Referee2_Response.tex

\noindent 
xxx  We agree. 
We changed the introduction to clearly state the contributions of the paper and \ldots  and the presentation of the paper significantly improved.  Thank you.
\medskip

\begin{itemize}
\item[2.] \textit{It is not necessary to go into such detail in the OBM review.}
\end{itemize}

\smallskip

\noindent 
We shortened and streamlined the OBM review, introducing the main results used in the paper, by combining the previous two sections---\S 2.1 (Overlapping Batch Means) and \S 3.1 (Variably Overlapping Batches)---along with other explanations on OBM that appeared later in the paper. 
The combined review is in the background section, \S 2.1 (Batch Means). 
To improve readability, we created two subsections: \S 2.1.1 (Nonoverlapping Batch Means) and \S 2.1.2 (Overlapping Batch Means). 
While we removed a number of equations and shortened the text, we kept some explanations that are necessary for a reader who is not familiar with OBM (e.g., readers with an optimization focus) to follow the main ideas. 
These basic explanations appear, for instance, in the first paragraph of \S 2.1.2 (Overlapping Batch Means).  
We hope to be able to reach a broader audience this way. 
As a result of these changes, 
\begin{itemize}
	\item[(i)]  instead of former equations (2) and (5), both on overlapping batch means variance estimator, we have only one equation (2);

@@

	\item[(ii)]  the confidence interval estimation using OBM is discussed only once;

	\item[(iii)] the former equation (3) is removed;

	\item[(iv)] the discussion that appeared at the beginning of \S 4.4 (previously: Variance Reduction, now: Lower Variances) is moved to this review section.\medskip

{\it Note:} While making the changes, we realized that `$\alpha$' was used in the current equation (3) and the statement of Theorem 2 as well as the ($1-\alpha$)-level confidence intervals.  
Therefore, we changed `$\alpha$' in equation (3) and Theorem 2 to `$\beta_{\bar{\gamma}}$'. 
\end{itemize}

\noi
With these changes, the OBM review became more streamlined, focusing on the main results used later in the paper. 
Thank you for this suggestion.

### 2014-07-05 omrp_lb_rev.tex
Various changes
1. Updated the conclusions,
2. Restructured Comp. Efficiency section, and
3. Updated the captions in Comp. Efficiency Section
4. Updated Ref1 and Ref2 responses

The amount of asymptotic reduction in $\var{\vartg{\ybb}}$ for a given $0\leq \gamma< m$ relative its nonoverlapping counterpart 
%$\var{\vtm{\ybb}}$ 
depends on the asymptotic ratio of $\gamma/m$, which we denote by $\gammab$. 
The maximally-overlapping case ($\gamma = 1$) corresponds to $\gammab=0$, and the nonoverlapping case corresponds to $\gammab=1$. 
\cite{damerdji1994strong,damerdji1995mean} studied batching structures that guarantee almost sure and mean-square consistency of $\vartg{\ybb}$ in the nonoverlapping and maximally-overlapping cases. 
Under appropriate conditions, these results, along with those of \cite{Meketon1984} and \cite{Welch1987}, establish that 
\begin{equation}
\label{eq:varvarred}
\frac{n}{m}\var{n \vartg{\ybb}} \rightarrow \beta_{\gammab} \sigma^4
\end{equation}
as the batch size $m$ and the number of batches $n/m$ tend to infinity. 
In (\ref{eq:varvarred}), $\beta_1 = 2$ for nonoverlapping batches, $\beta_0 = \frac{4}{3}$ for maximally overlapping batches, and more generally, $\beta_{\gammab}$ is twice the value of 
\begin{equation} \label{eq:var_reduct_formula}
	\gammab \left( 1 + 2 \sum_{j=1}^{\cogb-1} (1-j\gammab)^2 \right),
\end{equation}
where $\lceil \cdot \rceil$ denotes the smallest integer that is greater than or equal to its argument.
If $\ogb$ is an integer, (\ref{eq:var_reduct_formula}) simplifies to $\frac{2+\gammab^2}{3}$.
This means that the maximally-overlapping variance estimator has a variance that is two-thirds ($66.67\%$) of the nonoverlapping estimator.
As the amount of overlap decreases, the variance increases. 
When only half of the observations overlap ($\gammab = 1/2$ or $\gamma=m/4$), for instance, the variance is 75\% of the nonoverlapping counterpart.

A $(1-\alpha)$-level approximate CI on the mean can be formed using overlapping batches by $\ybb \pm t_{d_{\gamma}(k-1),\alpha/2}$ $\sqrt{\vartg{\ybb}}$, where the degrees of freedom increase $d_{\gamma}$ is the multiplicative inverse of (\ref{eq:var_reduct_formula}) \citep{Welch1987}.  
If $\ogb$ is an integer, $d_{\gamma} = \frac{3}{2+\gammab^2}$.  For the maximally-overlapping batch means, the degrees of freedom is $d_{\gamma} = \frac{3}{2}$.

@@

The internal optimization in the batches in (\ref{eq:gbar}) makes a straightforward statistical analysis of the behavior of the estimators difficult.  
To overcome this problem, we introduce the following unbiased optimality gap estimators:
\begin{align}
	\db_{lj} & = \frac{1}{m_l} \sum_{i=1}^{m_l} \left[ f(\xh,\xit^{\gamma_l(j-1)+i}) - f(\xs,\xit^{\gamma_l(j-1)+i}) \right],\ \ \ j = 1, 2, \dots, \nbl, \label{eq:dbar} \\
	\dbb_l & = \frac{1}{n_l} \sum_{i=1}^{n_l} \left[ f(\xh,\xiti) - f(\xs,\xiti) \right], \label{eq:dbb} \\
	VD_l & = \frac{1}{\left( \tfrac{n_l}{m_l} - 1 \right) \left(\nbl\right)} \sum_{j=1}^{\nbl} (\db_{lj} - \dbb_l)^2. \label{eq:vd}
\end{align}
These are essentially the same as the overlapping batches estimators in \S \ref{ssec:overlap} with $y^i = f(\xh,\xiti) - f(\xs,\xiti)$.  
These are also defined identically to the original estimators (\ref{eq:gbar})--(\ref{eq:vg}) with the exception that $\xs_j$ from (\ref{eq:gbar}) and (\ref{eq:gbb}) is replaced by an optimal solution $\xs$ in (\ref{eq:dbar}) and (\ref{eq:dbb}).  
With $\xh$ and $\xs$ fixed, estimators (\ref{eq:dbar})--(\ref{eq:vd}) have the same statistical properties as overlapping batches estimators in \S \ref{ssec:overlap}.  
The optimal solution $\xs$ is not known; however, the estimators (\ref{eq:dbar})--(\ref{eq:vd}) are used only to show convergence of (\ref{eq:gbar})--(\ref{eq:vg}) and not necessary for carrying out the OMRP algorithm.  
Throughout the rest of this section, we assume that the optimality gap estimators (\ref{eq:gbar})--(\ref{eq:vg}) and their nonoptimized counterparts (\ref{eq:dbar})--(\ref{eq:vd}) use identical samples and values of $n_
l$, $m_l$, and $\gamma_l$.  \smallskip

@@

\subsection{Computational Efficiency}
\label{ssec:compeff}

\begin{figure}[htb!]
	\centering
	\subfloat[][]{
	   \label{fig:efficiency_theory}
	   \includegraphics[width=.5\textwidth]{images/comp_efficiency_theory}}
	\subfloat[][]{
	   \label{fig:efficiency_nv}
	   \includegraphics[width=.5\textwidth]{images/nv_comp_efficiency}}
		\caption{
		Theoretical computational efficiency results for overlapping batches (left) and a comparison with measured computational efficiency for the Newsvendor problem using insertion sort (right).
		Both graphs are scaled so that nonoverlapping batches produces a computational efficiency of 1.
		}
\label{fig:efficiency}
\end{figure}

The increased computational workload from solving overlapping batches raises the question of the computational efficiency of this method to estimate $VG$.
In this setting, computational efficiency is defined as $[T(\gammab)\var{VG}]^{-1}$, where $T(\gammab)$ is the computational time spent to generate $VG$ for a given value of $\gammab$.
Because the variance of overlapping estimator relative to the nonoverlapping (MRP) variant is known from (\ref{eq:var_reduct_formula}), we need only examine the computational time, which can be decomposed as $T(\gammab) = t(\gammab)\left(\nb\right)$, where $t(\gammab)$ is the average computation time per batch, multiplied by the number of batches.
Warm-starting an algorithm may be used to decrease the computation time per batch $t(\gammab)$ for $\gammab < 1$, leading to a possible increase in computational efficiency.

To investigate computational efficiency, we assumed that computation time per batch can be approximated as $t(\gammab) \propto \gammab^p$ for some $p \geq 0$.
Results for various values of $p$ are shown in Figure \ref{fig:efficiency_theory}. 
This figure is constructed using $m = 1000$, but the results do not change substantially with different values of $m$.
Efficiency gains increase with larger exponent $p$, but are positive even for small values of $p$ when $\gammab$ is sufficiently close to 1.
At $p=1$, i.e., for problems where computational time is proportional to the number of samples replaced in the batch, computational efficiency is maximized with maximally overlapping batches of \cite{Meketon1984}.
When $p$ is $0.71$, on the other hand, computational efficiency is observed for values of $\gammab$ from slightly lower than $0.3$ to $1$, and efficiency is maximized at a relatively flat region between $\gammab=0.5$ and $\gammab=0.8$.

@@

We tested this theoretical relationship using the newsvendor problem solved with insertion sort.
Insertion sort is an efficient sorting algorithm when the values are already almost sorted. 
Therefore, as overlapping increases, insertion sort results in a decreased solution time per batch for the NVD problem. 
To test computational efficiency, we used $m = 1000$ with $10,000$ replications.
The average computation time per batch was found empirically to be well approximated ($R^2 = 0.994$) by power law with $p = 0.71$.
The empirically measured computational efficiency is compared to the theoretical result in Figure \ref{fig:efficiency_nv}, where the theoretical computational efficiency, shown in solid line, is repeated from Figure \ref{fig:efficiency_theory}.
It appears that empirical observations match with the theoretical efficiency calculations well.

In cases where the exponent $p$ can be estimated, the above theoretical method may be used to determine a reasonable value for $\gammab$.
When warm-starting can be used to solve the batch sampling problems, we recommend the users to perform preliminary computations to determine how computational time per batch changes as $\gammab$ changes. 
We can see in Figure \ref{fig:efficiency} that for some values of $p$, efficiency gain is nearly flat over a region of $\gammab$.
In this case, we recommend choosing a $\gammab$ near the right side of this region.
This choice keeps efficiency high while decreasing computation time and preventing large decreases in efficiency in the event that $p$ was overestimated.
For example, in the case of NVD, we recommend the use of value $\gammab=0.8$ as this value provides a high computational efficiency with minimal overall computation time across values of $\gammab$ with high efficiency.

\begin{figure}[htb!]
	\centering
	\subfloat[][]{
	   \label{fig:comp_work_comparison_theory}
	   \includegraphics[width=.5\textwidth]{images/work_comparison}}
	\subfloat[][]{
	   \label{fig:comp_work_comparison_nv}
	   \includegraphics[width=.5\textwidth]{images/nv_work_comparison}}
		\caption{
		(Left) A theoretical comparison of the computational work required to get equivalent variance reduction between OMRP and MRP when the number of batches are increased.
		(Right) Measured computational work for NVD compared with the theoretical results.
		}
\label{fig:comp_work_comparison}
\end{figure}

A final question on computational efficiency would be whether OMRP is more efficient than nonoverlapping MRP with a larger number of batches.
When generating observations is cheap, one could potentially increase the number of batches in MRP to obtain estimators with lower variances. 
In the nonoverlapping case, $\var{VG} \propto \frac{1}{k}$, so decreasing $\var{VG}$ to the level of maximally overlapping batches requires a $50\%$ increase in the number of batches, and thus a $50\%$ increase in computation time.
Continuing on in this way for various values of $\gammab$, we obtain the solid black curve in Figure \ref{fig:comp_work_comparison} using $m=1000$. 
This curve shows the increase in computational effort due to an increased number of batches in MRP in order to obtain the same variance level for OMRP across $\gammab$.
If the work increase by OMRP is less than this solid black curve, then OMRP is computationally more efficient than increasing the number of batches in MRP. 
Again, assuming average computation time per batch can be represented as $T(\gammab) \propto \gammab^p$, Figure \ref{fig:comp_work_comparison} compares the increase in computational work for OMRP for several values of $p$ with the work increase for MRP to get an equivalent decrease in variance.
We again see that OMRP is advantageous for larger values of $p$.
For small values of $p$, some efficiency gain is observed at $\gammab$ values close to 1.

With these considerations, for situations where $p > 0$ but its size cannot be readily measured or estimated, we recommend selecting a large value of $\gammab$ such as $0.9$.
This allows for some increase in efficiency while remaining robust in the event that $p$ is smaller than expected.

@@

We have applied the overlapping batch means method from simulation output analysis \citep{Meketon1984,Song1992,Welch1987} to assessing solution quality in stochastic programming, in particular, to Monte Carlo simulation-based estimators of optimality gaps \citep{Mak1999}. 
We have provided conditions under which the resulting point estimator of the optimality gap and its associated variance estimator are consistent, the same variance reduction can be achieved, and asymptotically valid CIs can be formed. 
Empirical results with small sample sizes indicate that the asymptotic reductions in variance of the variance estimator show a similar decrease in OMRP, while bias and coverage probability remain unaffected.

>>>>

The amount of asymptotic reduction in $\var{\vartg{\ybb}}$ for a given $0\leq \gamma< m$ relative its nonoverlapping counterpart 
%$\var{\vtm{\ybb}}$ 
depends on the asymptotic ratio of $\gamma/m$, which we denote by $\gammab$. 
The maximally-overlapping case ($\gamma = 1$) corresponds to $\gammab=0$, and the nonoverlapping case corresponds to $\gammab=1$. 
\cite{damerdji1994strong,damerdji1995mean} studied batching structures that guarantee almost sure and mean-square consistency of $\vartg{\ybb}$ in the nonoverlapping and maximally-overlapping cases. 
Under appropriate conditions, these results, along with those of \cite{Meketon1984} and \cite{Welch1987}, establish that 
\begin{equation}
\label{eq:varvarred}
\frac{n}{m}\var{n \vartg{\ybb}} \rightarrow \beta_{\gammab} \sigma^4
\end{equation}
as the batch size $m$ and the number of batches $n/m$ tend to infinity. 
In (\ref{eq:varvarred}), $\beta_1 = 2$ for nonoverlapping batches, $\beta_0 = \frac{4}{3}$ for maximally overlapping batches, and more generally, $\beta_{\gammab}$ is twice the value of 
\begin{equation} \label{eq:var_reduct_formula}
	\gammab \left( 1 + 2 \sum_{j=1}^{\cogb-1} (1-j\gammab)^2 \right),
\end{equation}
where $\lceil \cdot \rceil$ denotes the smallest integer that is greater than or equal to its argument.
If $\ogb$ is an integer, (\ref{eq:var_reduct_formula}) simplifies to $\frac{2+\gammab^2}{3}$.
This means that the maximally-overlapping variance estimator has a variance that is two-thirds ($66.67\%$) of the nonoverlapping estimator.
As the amount of overlap decreases, the variance increases. 
When only half of the observations overlap ($\gammab = 1/2$ or $\gamma=m/2$), for instance, the variance is 75\% of the nonoverlapping counterpart.

A $(1-\alpha)$-level approximate CI on the mean can be formed using overlapping batches by $\ybb \pm t_{d_{\gammab}(k-1),\alpha/2}$ $\sqrt{\vartg{\ybb}}$, where the degrees of freedom increase $d_{\gammab}$ is the multiplicative inverse of (\ref{eq:var_reduct_formula}) \citep{Welch1987}.  
%If $\ogb$ is an integer, $d_{\gammab} = \frac{3}{2+\gammab^2}$.  
For the maximally-overlapping batch means, the degrees of freedom is $d_{\gammab} = \frac{3}{2}$.

@@

The internal optimization in the batches in (\ref{eq:gbar}) makes a straightforward statistical analysis of the behavior of the estimators difficult.  
To overcome this problem, we introduce the following unbiased optimality gap estimators:
\begin{align}
	\db_{lj} & = \frac{1}{m_l} \sum_{i=1}^{m_l} \left[ f(\xh,\xit^{\gamma_l(j-1)+i}) - f(\xs,\xit^{\gamma_l(j-1)+i}) \right],\ \ \ j = 1, 2, \dots, \nbl, \label{eq:dbar} \\
	\dbb_l & = \frac{1}{n_l} \sum_{i=1}^{n_l} \left[ f(\xh,\xiti) - f(\xs,\xiti) \right], \label{eq:dbb} \\
	VD_l & = \frac{1}{\left( \tfrac{n_l}{m_l} - 1 \right) \left(\nbl\right)} \sum_{j=1}^{\nbl} (\db_{lj} - \dbb_l)^2. \label{eq:vd}
\end{align}
These are essentially the same as the overlapping batches estimators in \S \ref{ssec:overlap} with $y^i = f(\xh,\xiti) - f(\xs,\xiti)$.  
These are also defined identically to the original estimators (\ref{eq:gbar})--(\ref{eq:vg}) with the exception that $\xs_j$ from (\ref{eq:gbar}) and (\ref{eq:gbb}) is replaced by the optimal solution $\xs$ in (\ref{eq:dbar}) and (\ref{eq:dbb}).  
With $\xh$ and $\xs$ fixed, estimators (\ref{eq:dbar})--(\ref{eq:vd}) have the same statistical properties as overlapping batches estimators in \S \ref{ssec:overlap}.  
The optimal solution $\xs$ is not known; however, the estimators (\ref{eq:dbar})--(\ref{eq:vd}) are used only to show convergence of (\ref{eq:gbar})--(\ref{eq:vg}) and not necessary for carrying out the OMRP algorithm.  
Throughout the rest of this section, we assume that the optimality gap estimators (\ref{eq:gbar})--(\ref{eq:vg}) and their nonoptimized counterparts (\ref{eq:dbar})--(\ref{eq:vd}) use identical samples and values of $n_
l$, $m_l$, and $\gamma_l$.  \smallskip

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Computational Efficiency}
\label{ssec:compeff}

The increased computational workload from solving overlapping batches raises the question of the computational efficiency of this method to estimate $VG$.
In this setting, computational efficiency can be defined as $[T(\gammab)\var{VG}]^{-1}$, where $T(\gammab)$ is the computational time spent to generate $VG$ for a given value of $\gammab$.
Because the variance of the overlapping variance estimator relative to the nonoverlapping (MRP) variant is known from (\ref{eq:var_reduct_formula}), we need only examine the computational time, which can be decomposed as $T(\gammab) = t(\gammab)\left(\nb\right)$. 
Here, $t(\gammab)$ is the average computation time per batch, which is multiplied by the number of batches.
Warm-starting an algorithm may be used to decrease the computation time per batch for $\gammab < 1$, leading to a possible increase in computational efficiency.

To investigate computational efficiency, we assumed that computation time per batch can be approximated as $t(\gammab) \propto \gammab^p$ for some $p \geq 0$.
Results for various values of $p$ are shown in Figure \ref{fig:efficiency}(a). 
This figure is constructed using $m = 1,000$, but the results do not change substantially with different values of $m$.
The graph is also scaled so that nonoverlapping batches produces a computational efficiency of 1. 
Efficiency gains increase with larger exponent $p$ but are positive even for small values of $p$ when $\gammab$ is sufficiently close to 1.
At $p=1$, i.e., for problems where computational time is proportional to the number of samples replaced in the batch, computational efficiency is maximized with maximally overlapping batches of \cite{Meketon1984}.
When $p=0.9$, computational efficiency is maximized at around $\gammab=0.3$.
When $p$ is $0.71$, on the other hand, computational efficiency is observed for values of $\gammab$ from slightly lower than $0.3$ to $1$, and efficiency is maximized at a relatively flat region between $\gammab=0.5$ and $\gammab=0.8$.

\begin{figure}[thb!]
	\centering
	\subfloat[][Theoretical]{
	   \label{fig:efficiency_theory}
	   \includegraphics[width=.5\textwidth]{images/comp_efficiency_theory}}
	\subfloat[][Empirical for NVD]{
	   \label{fig:efficiency_nv}
	   \includegraphics[width=.5\textwidth]{images/nv_comp_efficiency}}
		\caption{
		Computational efficiency of the OMRP variance estimator relative to the (nonoverlaping) MRP variance estimator measured
		(a) theoretically and
		(b) empirically for NVD using insertion sort
%		(graphs are scaled so that nonoverlapping batches produces a computational efficiency of 1)
		}
\label{fig:efficiency}
\end{figure}

@@

We tested this theoretical relationship using the newsvendor problem solved with insertion sort.
Insertion sort is an efficient algorithm when the values are already largely sorted. 
Therefore, as overlapping increases, insertion sort results in a decreased solution time per batch for the NVD problem. 
To test computational efficiency, we used $m = 1,000$ with $10,000$ runs.
The average computation time per batch was found empirically to be well approximated ($R^2 = 0.994$) by the power law with $p = 0.71$.
The empirically measured computational efficiency is compared to the theoretical result in Figure \ref{fig:efficiency}(b), where the theoretical computational efficiency, shown in solid line, is repeated from Figure \ref{fig:efficiency}(a).
It appears that empirical observations match with the theoretical efficiency calculations well.

Another related question on computational efficiency would be whether OMRP is more efficient than nonoverlapping MRP with a larger number of batches.
When generating observations is cheap, one could alternatively increase the number of batches in MRP to obtain estimators with lower variances. 
In the nonoverlapping case, $\var{VG} \propto \frac{1}{k}$; so, decreasing $\var{VG}$ to the level of maximally overlapping batches of \cite{Meketon1984} requires a $50\%$ increase in the number of batches, and thus a $50\%$ increase in computation time.
Continuing on in this way for various values of $\gammab$, we obtain the solid black curve in Figure \ref{fig:comp_work_comparison}. 
This curve shows the increase in computational effort due to an increased number of batches in MRP in order to obtain the same variance for OMRP across $\gammab$.
If the work increase by OMRP is less than this solid black curve, then OMRP is computationally more efficient than increasing the number of batches in MRP. 
Again, assuming average computation time per batch can be represented as $T(\gammab) \propto \gammab^p$, Figure \ref{fig:comp_work_comparison}(a) compares the increase in computational work for OMRP for several values of $p$ to the work increase for MRP to get an equivalent decrease in variance.
We again see that OMRP is advantageous for larger values of $p$.
For small values of $p$, some efficiency gain is observed at $\gammab$ values close to 1.
We again tested this empirically on NVD using insertion sort with $m=1,000$, average of $10,000$ independent runs. 
The results, shown in Figure \ref{fig:comp_work_comparison}(b), again appear to match the theory well.

\begin{figure}[htb!]
	\centering
	\subfloat[][Theoretical]{
	   \label{fig:comp_work_comparison_theory}
	   \includegraphics[width=.5\textwidth]{images/work_comparison}}
	\subfloat[][Empirical for NVD]{
	   \label{fig:comp_work_comparison_nv}
	   \includegraphics[width=.5\textwidth]{images/nv_work_comparison}}
		\caption{
		Comparison of computational work required to get equivalent variance between MRP variance estimator with increased number of batches and OMRP variance estimator measured
		(a) theoretically and 
		(b) empirically for NVD using insertion sort
%		(Left) A theoretical comparison of the computational work required to get equivalent variance reduction between OMRP and MRP when the number of batches are increased.
%		(Right) Measured computational work for NVD compared with the theoretical results.
		}
\label{fig:comp_work_comparison}
\end{figure}

In cases where the exponent $p$ can be estimated, the above theoretical method may be used to determine a reasonable value for $\gammab$.
When warm-starting can be used to solve the batch sampling problems, we recommend the users to perform preliminary computations to determine how computational time per batch changes as $\gammab$ changes. 
We can see in Figure \ref{fig:efficiency} that for some values of $p$, efficiency gain is nearly flat over a region of $\gammab$.
In this case, we recommend choosing a $\gammab$ near the right side of this region.
This choice keeps efficiency high while decreasing computation time and preventing large decreases in efficiency in the event that $p$ was overestimated.
For example, in the case of NVD, we recommend the use of value $\gammab=0.8$ as this value provides high computational efficiency with minimal overall computation time across values of $\gammab$ with high efficiency. 
For situations where $p > 0$ but its size cannot be readily measured or estimated, we recommend selecting a large value of $\gammab$ such as $0.9$.
This allows for some increase in efficiency while remaining robust in the event that $p$ is smaller than expected.

@@

We have applied the overlapping batch means method from simulation output analysis \citep{Meketon1984,Song1992,Welch1987} to assessing solution quality in stochastic programming, in particular, to Monte Carlo simulation-based estimators of optimality gaps \citep{Mak1999}. 
We have provided conditions under which the resulting point estimator of the optimality gap and its associated variance estimator are consistent, the same lower variances of OMRP variance estimators relative to the nononverlapping MRP estimator can be achieved, and asymptotically valid CIs can be formed. 
Empirical results with small sample sizes indicate that the asymptotic reductions in variance of the variance estimator show a similar decrease in OMRP, while bias and coverage probability remain unaffected. 
We have examined the computational efficiency of OMRP by jointly considering increases in computation time and decreases in variance. 
Our results indicate that when warm-starting is effective, efficiency can be observed across a wide range of degree of overlap. 
When warm-starting is less effective, efficiency is observed for small degrees of overlap. 
We have demonstrated computational efficiency empirically solving the newsvendor problem with insertion sort.
Finally, we provided guidelines on selecting the degree of overlap based on our study of computational efficiency.
